\LARGE \bf Crecimiento económico y su relación con energías limpias en la región latinoamericana

### 2020-06-29 paper/main.tex
Add bibliography to paper file

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[spanish]{babel}

@@

This template, modified \cite{noauthor_why_2014} in MS Word 2003 and saved as ``Word 97-2003 \& 6.0/95 -- RTF'' for the PC, provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

\section{Metodología}

This template, modified \cite{noauthor_why_2014} in MS Word 2003 and saved as ``Word 97-2003 \& 6.0/95 -- RTF'' for the PC, provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[spanish]{babel}
%Bibliograpy
\usepackage{cite}

@@

This template, modified  in MS Word 2003 and saved as ``Word 97-2003 \& 6.0/95 -- RTF'' for the PC, provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

\section{Metodología }

Para la realización de este trabajo se decidió utilizar la metodología propuesta por Benjamín Fry \cite{fry_visualizing_2007}, este método consta de siete etapas:
\begin{itemize}
    \item \textbf{Recolectar datos: } 
    esta etapa consiste en buscar y obtener el conjunto de datos que vamos a utilizar, lo principal es conseguir fuentes de datos confiables, provenientes de otros estudios académicos o algún organismo oficial. En esta fase también se define como los datos recolectados van a ser accedidos por los usuarios.
    \item \textbf{Estructuración de los datos: } 
    una vez que se tienen datos confiables, hay que ordenarlos y categorizarlos para que estos adquieran una estructura bien definida. Si lo datos no están estructurados correctamente se dificulta el análisis y la visualización de los mismos.
    \item \textbf{Filtrar: }
    en este paso se debe decidir desde el punto de vista de la visualización cuales datos son útiles y aportan información relevante. Los datos que no se consideren relevantes son filtrados y no se toman en cuenta para la representación.
    \item \textbf{Extraer: }
    se procede a convertir, los datos relevantes, en variables que denoten los valores o cantidades que necesitamos analizar y mostrar a través de la visualización. Esto se logra al aplicar métodos estadísticos o de minería de datos que permitan colocar los datos en un contexto matemático.
    \item \textbf{Representar:}
    se escoge un modelo visual básico (Gráfico de barras, arboles, listas, entre otros), apropiado para una visualización clara y concisa de las variables creadas en el paso anterior.
    \item \textbf{Refinar: }
    trabajar de forma iterativa sobre el modelo visual básico con el fin de mejorarlo y que sea capaz de transmitir la información de forma clara, además debe ser atractivo y cautivador para el usuario.  
    \item \textbf{Interactuar: }
    agregar elementos que permitan al usuario la manipulación y el control de las características que son visibles. Los usuarios deben poder seleccionar diferentes rangos de datos e intervalos de tiempo que le permitan analizar la información presentada desde múltiples puntos de vista.  
\end{itemize}
